---
name: Bas
surname: van der Burg
position: "Student"
phone: 06 28304767
profilepic: bas.jpeg
www: bburg2.github.io
email: "Bas.vanderburg@student.hu.nl"
github: bburg2
linkedin: soontm
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "I am a life science student who specializes in data analysis."
output: vitae::twentyseconds
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

# Some stuff about me

 * I have done half a year of programming before life science in which I learned java and SQL.
 * I have been swimming my whole life and have experience in rescue swimming and scuba diving. 
 * I have experience working with DNA-sequencing with the help of the minION.
 * I can work with the agile workflow and work with a scrum board

# Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Bachelor", "2018-2023", "Hogeschool Utrecht", "Utrecht, Netherlands",
  "Minor in data science", "2022", "Hogeschool Utrecht", "Utrecht, Netherlands",
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# lab techniques

```{r}
tribble(
  ~ Desc, ~ Year, ~ Institution, ~ Where,
  "DNA-sequencing", "2018-2023", "Hogeschool Utrecht", "Utrecht, Netherlands",
  "PCR", "2018-2023", "Hogeschool Utrecht", "Utrecht, Netherlands",
  "Gel-electroforese", "2018-2023", "Hogeschool Utrecht", "Utrecht, Netherlands",
) %>% 
  detailed_entries(Desc, Year, Institution, Where)
```

# Languages 

```{r}
tribble(
  ~ Desc, ~ What, ~ State,
  "Native language", "Dutch", "Fluent", 
  "Learned language", "English", "Good",
  "Learned language (see website)", "R","Novice",
  "Learned language (see website)", "SQL","Novice",
  "Learned language (see website)", "bash","Novice",
  "Learned language (see website)", "java","Beginner",
) %>% 
  detailed_entries(Desc, What, State)
```

# Work Experience

```{r}
tribble(
  ~ Desc, ~ Year, ~ Where,
  "Working at the warehouse restocking the supermarket", "2015-2016", "Albert heijn",
  "Data managment for a company building a website", "2018","Dr. Oetker",
) %>% 
  detailed_entries(Desc, Year, Where)
```

